Taking uncertainty in word embedding bias estimation seriously: a  bayesian approach

### 2021-07-13 paperDraft/paperDraft.Rmd
small style changes

---
title: "A bayesian method of cosine-based word2vec bias"
author: "Alicja Dobrzeniecka and Rafal Urbaniak"
output:
  pdf_document:
    number_sections: true
    df_print: kable 
    keep_tex: true
    includes:
      in_header:
        - Rafal_latex6.sty
classoption: dvipsnames,enabledeprecatedfontcommands
fontsize: 10pt
documentclass: scrartcl
urlcolor: blue
bibliography: cosineReferences1.bib
csl: apa-5th-edition.csl
---

@@

The attributes are taken from different sources, there is no principled justification for their choice. From our analysis it will become clear that the list is rather uneven.

There is no mention of methodology for deciding on the number of attributes necessary to decide a hypothesis on the given size of dataset. There are however some ways to estimate sample sizes needed to ensure that the results are meaningful if the effect is present. Our research will show that the numbers used
are rather insufficient.

\textbf{2. No attention to distributions and details}

The authors use the mean of mean average cosine similarities to measure multi-class similarity between protected word and harmful stereotypes. They average the results until they obtain one final value to represent the mean cosine distance between protected word from a given class and the attributes of that class. If one takes closer look at the individual values that are taken for the calculations it turns out that there is a bunch of outliers and surprisingly dissimilar words. We approach this issue by providing the tables and visualizations of individual cosine distances to make sure that we obtain a proper insight into the data.

@@

grid.arrange(t1d2Plot,t2d2Plot, t3d2Plot, t4d2Plot, ncol=2,   
      top = textGrob("Different means, sd=0.05",gp=gpar(fontsize=15,font=1)))
```
\normalsize

@@

Let's analyse dataset-level coeﬀicients together with their highest posterior density intervals, for the three topic groups. We compare the results for two different word emebeddings corpuses - RedditL2 and GoogleNews. For Google embeddings the HPDIs
for all class coeﬀicients (associated, different, human, and none) include zero. This can lead one to a conclusion that the impact for associated, different, human and neutral attribute is, when averaged, quite similar. This indicates how including the uncertainty
may change the use and interpretation of Manzini, Lim, Tsvetkov, & Black (2019) MAC metric. It seems that if one focuses only on differences between means of means, this is too simplistic.

In the case of Reddit word embeddings the situation is similar although HPDI is below 0 for Gender class when looking at `associated` and `different` mean coeﬀicient. This can suggest that there is indeed slightly stronger impact of these attributes. One should also notice how in general associated and different coeﬀicients have quite similar HPDI range, the highest observed absolute difference is equal to only 0.1. This suggests that again the impact of associated attributes and different ones is not clear at first sight. Finally, let’s compare the HPDIs for Reddit and Reddit debiased datasets. For Religion and Race dataset there is a minor shift (in absolute values the highest change is equal to approximately 0.1) of the mean coeﬀicients towards zero. However for Gender dataset there is no significant change. This is of course a general look at the group coeﬀicients. Let’s now analyze the individual words.

>>>>

---
title: "Taking uncertainty in word embedding bias estimation seriously: a  bayesian approach"
author: "Alicja Dobrzeniecka and Rafal Urbaniak"
output:
  pdf_document:
    number_sections: true
    df_print: kable 
    keep_tex: true
    includes:
      in_header:
        - Rafal_latex6.sty
classoption: dvipsnames,enabledeprecatedfontcommands
fontsize: 10pt
documentclass: scrartcl
urlcolor: blue
bibliography: [../references/bibliographyRafal.bib]
csl: [../references/apa-6th-edition.csl] 
---

@@

Such approaches, however, are not unproblematic. For one thing, the word lists are often fairly small, and the papers do not discuss the impact of word list sizes on the uncertainty involved. 
 We will argue that the word list sizes are too small to yield strong conclusions, given the variances involved.

 Crucially, these approaches use  means of mean average cosine similarities to measure  similarity between protected word and harmful stereotypes. If one takes  a closer look at the individual values that are taken for the calculations, it turns out that there are quite a few  outliers and surprisingly dissimilar words. This issue will become transparent when  we inspect  the visualizations of individual cosine distances, following the idea that one of the first step to understand data is to look at it.

@@

grid.arrange(t1d2Plot,t2d2Plot, t3d2Plot, t4d2Plot, ncol=2,   
      top = textGrob("different means, sd=0.05",gp=gpar(fontsize=15,font=1)))
```
\normalsize

@@

Let's analyse dataset-level coefficients together with their highest posterior density intervals, for the three topic groups. We compare the results for two different word emebeddings corpuses - RedditL2 and GoogleNews. For Google embeddings the HPDIs
for all class coefficients (associated, different, human, and none) include zero. This can lead one to a conclusion that the impact for associated, different, human and neutral attribute is, when averaged, quite similar. This indicates how including the uncertainty
may change the use and interpretation of Manzini, Lim, Tsvetkov, & Black (2019) MAC metric. It seems that if one focuses only on differences between means of means, this is too simplistic.

In the case of Reddit word embeddings the situation is similar although HPDI is below 0 for Gender class when looking at `associated` and `different` mean coefficient. This can suggest that there is indeed slightly stronger impact of these attributes. One should also notice how in general associated and different coefficients have quite similar HPDI range, the highest observed absolute difference is equal to only 0.1. This suggests that again the impact of associated attributes and different ones is not clear at first sight. Finally, let’s compare the HPDIs for Reddit and Reddit debiased datasets. For Religion and Race dataset there is a minor shift (in absolute values the highest change is equal to approximately 0.1) of the mean coefficients towards zero. However for Gender dataset there is no significant change. This is of course a general look at the group coefficients. Let’s now analyze the individual words.
